title: "Human Activity Recognition"
author: "Haftom"
date: "October 19, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Background
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).


####Data
The training data for this project are available here:
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv
The test data are available here:
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv
The data for this project come from this source: http://groupware.les.inf.puc-rio.br/har. If you use the document you create for this class for any purpose please cite them as they have been very generous in allowing their data to be used for this kind of assignment.

####Getting ready
```{r echo = TRUE}
rm(list=ls())
require(tree)
require(ggplot2)
require(caTools)
require(caret)
require(dplyr)
library(randomForest)
library(e1071)
library(gbm)
set.seed(12345)
```

####Getting Data
```{r echo = TRUE}
# download the datasets
train<- read.csv("http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv")
sample<-sample.split(train$classe,SplitRatio = 0.70)
train_data<-subset(train,sample==TRUE)
test_data<-subset(train,sample==FALSE)
```


####Handling NA's

So much NA's would be removed using The Near Zero variance (NZV).
```{r }
#Checking for NA
table(is.na(train))
# remove variables with Nearly Zero Variance
remove1 <- nearZeroVar(train_data)
Train_c <- train_data[, -remove1]
Test_c  <- test_data[, -remove1]
dim(Train_c)
```


Removing columns that contains so much NA's
```{r }
remove2<- sapply(Train_c, function(x) mean(is.na(x))) > 0.95
Train_final <- Train_c[, remove2==FALSE]
Test_final  <- Test_c[,remove2==FALSE]

#Removing columns that are not required for modleing  
Train_final <- Train_final[, -(1:5)]
Test_final <- Test_final[, -(1:5)]

dim(Train_final)
dim(Test_final)
```
####Model One (Tree)
```{r }
#Training 
tree_model<-tree(classe~.,data=Train_final)
summary(tree_model)
plot(tree_model)
text(tree_model,pretty=0)


#Testing 
predict_Tree <- predict(tree_model, newdata=Test_final, type="class")
conf_Tree <- confusionMatrix(predict_Tree, Test_final$classe)
conf_Tree

```

####Model two (Random Forest)
```{r }
#Modeling
forest_model<-randomForest(classe~.,Train_final,mtyr=54,importance=TRUE)
forest_model
plot(forest_model)
#Testing 
predict_forest <- predict(forest_model, newdata=Test_final, type="class")
conf_forest <- confusionMatrix(predict_forest, Test_final$classe)
conf_forest
```
###Selecting a Modle
The accuracy of the tworegression modeling are
:
Decision Tree : 0.6978 

Random Forest : 0.9968 

Rondom Foreset is a good candidate

###Testing Data
```{r }
#Reading test data
real_test<-read.csv("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv")
#Predict test data
predict_test<- predict(forest_model,newdata=real_test,type="class")
head(predict_test)
```

